---
title: "Mapping Assignment"
author: "Brandan P. Buck"
date: "6 Apr 2019"
output: html_document
---

```{r, include=FALSE}

library(dplyr)
library(tidyverse)
library(leaflet)
library(sf)
library(maps)
library(mapdata)

sinking_data<- read_csv("C:/Users/bpbuc/Documents/Brandan's Stuff/_PhD Stuff/CILO II/project_data/sinking_data/japanese_merchant_fleet_sinkings_simplified_v4.csv")
sloc<- read_sf("C:/Users/bpbuc/Documents/Brandan's Stuff/_PhD Stuff/Cilo Wired-Hist 696/Project/Data/SLOC/Japanese SLOC WWII Complete.shp")

```

For this assignment I revisited the sinking data that I intend to use for the final project. I could not figure out how to use ggplot to build static maps since the data I am using already has geocoords. Also, since these events occur on the ocean, I don’t have a polygon shapefile to use as a baselayer in a choropleth map. Therefore, I made two leaf maps, one depicting mass causality events and the other the all sinking events from May 1945 until the end of the war, chief among them being the mining operation known as Operation Starvation.


```{r}

dms2dd <- function(x) {
  deg <- str_extract(x, "\\d+") %>% as.numeric()
  min <- str_extract(x, "-\\d+") %>% str_remove("-") %>% as.numeric()
  deg <- deg + min / 60
  dir <- str_extract(x, "\\w$")
  ifelse(dir %in% c("N", "E"),
         deg,
         -deg)}

sinking_data_dd<- sinking_data %>%
  mutate(lat= dms2dd(loc_1),
         lon= dms2dd(loc_2))

sinking_data$date<-as.Date(sinking_data$date, '%Y-%m-%d')

pal <- colorFactor(c("red", "green", "blue", "purple"), domain = c("aircraft", "mine", "submarine", "surface_craft"))

sinking_data_dd_42 <-sinking_data_dd %>%
  drop_na(cas_est) %>%
  filter(year == 1942) %>%
  filter(cas_est > 150)

sinking_data_dd_43 <-sinking_data_dd %>%
  drop_na(cas_est) %>%
  filter(year == 1943) %>%
  filter(cas_est > 150)

sinking_data_dd_44 <-sinking_data_dd %>%
  drop_na(cas_est) %>%
  filter(year == 1944) %>%
  filter(cas_est > 150)

sinking_data_dd_45 <-sinking_data_dd %>%
  drop_na(cas_est) %>%
  filter(year == 1945) %>%
  filter(cas_est > 150)

sinking_data_dd_cas_est<- sinking_data_dd %>% 
  drop_na(cas_est) %>%
  filter(cas_est > 150)

downfall_sub<- sinking_data_dd %>% 
  filter(date > "1945-04-30",
         sinking_agent == "submarine")

downfall_mine<- sinking_data_dd %>% 
  filter(date > "1945-04-30",
         sinking_agent == "mine")

downfall_aircraft<- sinking_data_dd %>% 
  filter(date > "1945-04-30",
         sinking_agent == "aircraft")

downfall<- sinking_data_dd %>% 
  filter(date > "1945-04-30")
```

This is an interactive map showing all vessels with known casualty estimates over 150. The circles have been colored to indicate sinking agent, blue for submarines, red for aircraft, green for mines and purple for surface craft. A cursory look at the data shows that sinkings via submarines constituted an overwhelming number of mass causality sinkings, 84 for subs, 19 for aircraft and two each for surface vessels and mines. These mass casualty events constituted 127,941 estimated causalities, 84% of the estimated loss of life during the Allied campaign to sink the Japanese merchant fleet.  

The largest number of these events clustered between the island of Luzon and Taiwan. Examining each event also shows at the number of mass causality events occurred during predominately in 1944. These high casualty events occurred because as the Japanese merchant fleet moved Allied POWs towards the interior of the empire and Japanese to the periphery to defend against Allied advances.  Several, isolated vessels were sunk my U.S. submarines in the waters south of Japan between Okinawa and Iwo Jima.

U.S. submarines sunk cluster of vessels which resulted in mass causality events south of the Japanese island of Kyushu and in the China Sea. These events occurred when only Allied submarines (almost exclusively U.S. boats) were capable of sinking Japanese merchant vessels in vicinity of the Japanese home islands. 

This amended dataset displays the same spatial pattern as the larger one, submarine sinkings constituted a war of attrition clustered in, but not limited to the South China Sea. Aircraft however mass their assaults on major ports and shipping nodes. Port facilities on the coral atoll of Truk and southern Taiwan depict this pattern prominently. Similarly, Allied submarines, as one may expect had the longest reach throughout war. Conversely, Allied aircraft played a larger role, later in the war as larger carrier groups won decisive battles with the Imperial Japanese Navy.


```{r}

leaflet() %>%
  addProviderTiles("Stamen.Watercolor", group = "map") %>%
  addCircleMarkers(data = sinking_data_dd_42,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   group = "1942",
                   popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " ")) %>%
  addCircleMarkers(data = sinking_data_dd_43,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   group = "1943",
                   popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " ")) %>% 
  addCircleMarkers(data = sinking_data_dd_44,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   group = "1944",
                   popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " ")) %>% 
  addCircleMarkers(data = sinking_data_dd_45,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   group = "1945",
                   popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " ")) %>% 
  addCircleMarkers(data = sinking_data_dd_cas_est,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   group = "1942 - 1945",
                   popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " ")) %>%
  addLegend(position = "bottomright",
            pal= pal, 
            values = c("aircraft", "mine", "submarine", "surface_craft"),
            title = "Sinking Agent") %>% 
  addLayersControl(
    baseGroups = c("1942","1943","1944","1945","1942 - 1945"))


```

This is a map of all ships sunk by U.S. Army Air Force air dropped mines, aircraft and subs from May 1945 until the end of the war. Exploring the map shows the various clusters of mining activity and the overwhelming density sinkings within the Kanmon straits. As one reduces the viewing scale, the extend of the mining campaign becomes apparent as seemingly every inlet saw a merchant ship sink. Smaller clusters of mining activity occurred on outside of the Port of Niigata, other port facilities on the western coast of the island of Honshu, and off the Korean coast.  

```{r}
leaflet() %>%
  addProviderTiles("Stamen.Watercolor", group = "map") %>%
  addMarkers(data = downfall_aircraft,
                   lng = ~lon,
                   lat = ~lat,
                   group = "Aircraft",
                   popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " "), 
                   clusterOptions = markerClusterOptions()) %>% 
  addMarkers(data = downfall_mine,
                   lng = ~lon,
                   lat = ~lat,
                   group = "Mine",
                   popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " "), 
                   clusterOptions = markerClusterOptions()) %>% 
  addMarkers(data = downfall_sub,
                   lng = ~lon,
                   lat = ~lat,
                   group = "Submarine",
                   popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " "), 
                   clusterOptions = markerClusterOptions()) %>% 
  addMarkers(data = downfall,
             lng = ~lon,
             lat = ~lat,
             group = "All Sinking Agents",
             popup = ~paste(name, cas_est, type, sinking_agent, date, sep = " "), 
             clusterOptions = markerClusterOptions()) %>%
  addLayersControl(
    baseGroups = c("Aircraft","Mine","Submarine", "All Sinking Agents"))

```

Below is an interactive map of the entire campaign. A larger view is available at:
http://brandanpbuck.com/unrestricted/unrestricted.html

```{r}
sinking_data_dd_41_42 <-sinking_data_dd %>%
  filter(year == 1941 |
         year == 1942) %>%
  filter(sinking_agent != "sabotage" |
         sinking_agent != "unknown")

sinking_data_dd_jan_jun_43 <-sinking_data_dd %>%
  filter(year == 1943) %>%
  filter(date < "1943-07-01") %>%
  filter(sinking_agent != "sabotage" |
           sinking_agent != "unknown")

sinking_data_dd_jul_dec_43 <-sinking_data_dd %>%
  filter(year == 1943) %>%
  filter(date > "1943-06-28") %>%
  filter(sinking_agent != "sabotage" |
           sinking_agent != "unknown")

sinking_data_dd_jan_mar_44 <-sinking_data_dd %>%
  filter(year == 1944) %>%
  filter(date < "1944-04-01") %>%
  filter (sinking_agent != "sabotage" |
            sinking_agent != "unknown")

sinking_data_dd_apr_sep_44 <-sinking_data_dd %>%
  filter(year == 1944) %>%
  filter(date < "1944-10-01") %>%
  filter(date > "1944-03-30") %>%
  filter (sinking_agent != "sabotage" |
            sinking_agent != "unknown")

sloc_closed_mar_44<- sloc %>%
  filter(Closed != "March 44")

sinking_data_dd_oct_dec_44 <-sinking_data_dd %>%
  filter(year == 1944) %>%
  filter(date > "1944-09-29",
         sinking_agent != "sabotage" |
           sinking_agent != "unknown")

sloc_closed_may_44<- sloc %>%
  filter(Closed != "March 44") %>%
  filter(Closed != "May 44")

sinking_data_dd_jan_apr_45 <-sinking_data_dd %>%
  filter(year == 1945) %>%
  filter(date < "1945-04-30",
         sinking_agent != "sabotage" |
           sinking_agent != "unknown")

sloc_closed_dec_44<- sloc %>%
  filter(Closed != "March 44") %>%
  filter(Closed != "May 44") %>%
  filter(Closed != "Dec 44")

sinking_data_dd_may_aug_45 <-sinking_data_dd %>%
  filter(year == 1945) %>%
  filter(date > "1945-04-30",
         sinking_agent != "sabotage" |
           sinking_agent != "unknown")

sloc_closed_may_45<- sloc %>%
  filter(Closed != "March 44") %>%
  filter(Closed != "May 44") %>%
  filter(Closed != "Dec 44") %>%
  filter(Closed != "May 45")


leaflet() %>%
  addProviderTiles("Stamen.Watercolor", group = "Map") %>%
  addProviderTiles("Esri.WorldImagery", group = "Imagery") %>%
  addCircleMarkers(data = sinking_data_dd_41_42,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   radius = 3,
                   popup = ~paste(name, type, sinking_agent, date, sep = ","),
                   group = "1. Dec, 1941- Dec, 1942") %>%
  addPolylines(data = sloc,
               color = "grey",
               group = "1. Dec, 1941- Dec, 1942") %>%
  addCircleMarkers(data = sinking_data_dd_jan_jun_43,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   radius = 3,
                   popup = ~paste(name, type, sinking_agent, date, sep = ","),
                   group = "2. Jan-Jun 1943") %>%
  addPolylines(data = sloc,
               color = "grey",
               group = "2. Jan-Jun 1943") %>%
  addCircleMarkers(data = sinking_data_dd_jul_dec_43,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   radius = 3,
                   popup = ~paste(name, type, sinking_agent, date, sep = ","),
                   group = "3. Jul - Dec, 1943") %>%
  addPolylines(data = sloc,
               color = "grey",
               group = "3. Jul - Dec, 1943") %>%
  addCircleMarkers(data = sinking_data_dd_jan_mar_44,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   radius = 3,
                   popup = ~paste(name, type, sinking_agent, date, sep = ","),
                   group = "4. Jan-Mar, 1944") %>%
  addPolylines(data = sloc,
               color = "grey",
               group = "4. Jan-Mar, 1944") %>%
  addCircleMarkers(data = sinking_data_dd_apr_sep_44,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   radius = 3,
                   popup = ~paste(name, type, sinking_agent, date, sep = ","),
                   group = "5. Apr-Sep, 1944") %>%
  addPolylines(data = sloc_closed_mar_44,
               color = "grey",
               group = "5. Apr-Sep, 1944") %>%
  addCircleMarkers(data = sinking_data_dd_oct_dec_44,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   radius = 3,
                   popup = ~paste(name, type, sinking_agent, date, sep = ","),
                   group = "6. Oct-Dec, 1944") %>%
  addPolylines(data = sloc_closed_may_44,
               color = "grey",
               group = "6. Oct-Dec, 1944") %>%
  addCircleMarkers(data = sinking_data_dd_jan_apr_45,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   radius = 3,
                   popup = ~paste(name, type, sinking_agent, date, sep = " "),
                   group = "7. Jan-Apr, 1945") %>%
  addPolylines(data = sloc_closed_dec_44,
               color = "grey",
               group = "7. Jan-Apr, 1945") %>%
  addCircleMarkers(data = sinking_data_dd_may_aug_45,
                   lng = ~lon,
                   lat = ~lat,
                   color = ~pal(sinking_agent),
                   radius = 3,
                   popup = ~paste(name, type, sinking_agent, date, sep = ","),
                   group = "8. May-Aug, 1945") %>%
  addPolylines(data = sloc_closed_may_45,
               color = "grey",
               group = "8. May-Aug, 1945") %>%
  addLegend(position = "bottomright",
            pal= pal,
            values = c("aircraft", "mine", "submarine", "surface_craft"),
            title = "Sinking Agent") %>%
  addLayersControl(
    baseGroups = c("Map", "Imagery"),
    overlayGroups = c("1. Dec, 1941- Dec, 1942","2. Jan-Jun 1943","3. Jul - Dec, 1943",
                      "4. Jan-Mar, 1944","5. Apr-Sep, 1944",
                      "6. Oct-Dec, 1944", "7. Jan-Apr, 1945", "8. May-Aug, 1945")) %>%
  addScaleBar(position = "bottomleft",
              options = scaleBarOptions(maxWidth = 250, imperial = TRUE,
                        metric = FALSE, updateWhenIdle = TRUE))

```



